


\exno[5]{Linear Programming Relaxations for weighted vertex cover}

\newcommand{\xIntgMin}{\bbv{x}\rmsc{\!I}}
\newcommand{\xRealMin}{\bbv{x}\rmsc{\!R}}
Given a fixed minimisation direction $\bbv{c}\in\mathbb{R}^n$ and
$m$ constraint specifiers $\bbv{a_j}\in\mathbb{R}^n$, $b_j\in\mathbb{R}$,
let $\xIntgMin\in\{0,1\}^n$ and $\xRealMin\in[0,1]^n$ be solutions to
the problem of minimising $\braKet{\bbv{c}}{\bbv{x}}$ under the constraint
$\braKet{\bbv{a}_j}{\bbv{x}}\geq b_j$, for integral and connected domain respectively.
Finding such solutions is the task of the ILP / LP linear-programming problems.

\subex{}\label{continuousrules}
$\xIntgMin$ can never solve the problem better than $\xRealMin$,
because $\{0,1\}^n \subset [0,1]^n$, so, if
$\braKet{\bbv{c}}{\xIntgMin} < \braKet{\bbv{c}}{\xRealMin}$ were the
case, then $\xIntgMin$ would also have to be the solution to the 
continuous minimisation problem, and $\xRealMin$ would \emph{not} be
the global minimum, in contradiction to the assumption.

\subex{}

Let $(V,E)$ a graph and $c\colon V \to \mathbb{R}$ a cost function. We seek the
minimal weighted vertex cover, i.e. a vertex subset $S \in \mathbb{P}(V)$ with the
property that all $(u,v)\in E$ intersect $S$, such that
\[
  \sum_{v\in S}c_v
\]
\newcommand{\xSharpI}{\bbv{x}^{\sharp}_{\mathrm{ILP}}}
\newcommand{\xSharp}{\bbv{x}^{\sharp}_{\mathrm{LP}}}
is minimal. This can equivalently be formulated as the search for a characteristic
vector $\xSharpI \in \{0,1\}^{|V|} \simeq (V\to \{0,1\}) \simeq \mathbb{P}(V)$ such that
\[
  \sum_{v\in V}c_v \cdot (\xSharpI)_v
    = \braKet{\bbv{c}}{\xSharpI}
\]
is minimal, while obeying
\[
  \sum_{v\in V} (\delta_{vu} + \delta_{vw})\cdot (\xSharpI)_v \geq 1  \qquad { \forall (u,w)\in E }
\]
where $\delta$ is the Kronecker symbol, indicating as 1 when an edge-end is in the sum.

Obtaining such a minimum is a task for ILP.
The result of the minimisation can be read as a vertex cover by letting
\newcommand{\covI}{S\rmsc{ILP}}
\newcommand{\cov}{S\rmsc{LP}}
\[
  { \covI = \{v\in V : (\xSharpI)_v = 1\} }.
\]

\subex{}
Unlike ILP, the continuous LP can not be expected to yield any components equal to
1 (or any other concrete number), which could be interpreted as clear membership or
not-membership in a vertex cover.
However, given a solution $\xSharp \in [0,1]^{|V|}$ to the specification above but
continuous domain, we can define
\[
  { \cov = \{v\in V : (\xSharp)_v \geq \tfrac12\} }.
\]
Claim: \emph{this still defines a vertex cover, and has a cost of at most a constant multiple
of the optimal solution $\xSharpI$}.
\\\\
Proof: we have required
\[
  \sum_{v\in V} (\delta_{vu} + \delta_{vw})\cdot (\xSharpI)_v \geq 1  \qquad { \forall (u,w)\in E }.
\]
For any edge $(u,w)$, this sum actually has only two nonzero terms:
\[
  (\xSharp)_u + (\xSharp)_w \geq 1.
\]
Because the $(\xSharp)_v$ are all in $[0,1]$, this will be fulfilled iff either
$(\xSharp)_u \geq \tfrac12$ or $(\xSharp)_w \geq \tfrac12$ (or both),
meaning that either vertex $u$ or $w$ is in $\cov$.
Holding for all edges, this implies that $\cov$ is a vertex cover.

\newcommand{\cost}{\mathrm{cost}}
Furthermore, the cost of this solution is bounded from above by $2\cdot\cost(\covI)$:
\[
  \cost(\cov)
    = \sum_{v\in \cov} c_v
    = \quad \sum_{\mathrlap{v\in V : (\xSharp)_v\geq\tfrac12}} c_v
    \leq\quad 
        \sum_{\mathrlap{v\in V : (\xSharp)_v<\tfrac12}}
            {(\text{anything }\geq 0)}
        + \sum_{\mathrlap{v\in V : (\xSharp)_v\geq\tfrac12}}
             c_v \cdot{(\text{anything }\geq 1)}
    \leq\quad  \sum_{v\in V} c_v \cdot 2\cdot (\xSharp)_v
\]
because $2\cdot (\xSharp)_v$ is always $\geq 0$, and even $\geq 1$
whenever $(\xSharp)_v\geq\tfrac12$. Now,
\[
  \sum_{v\in V}c_v \cdot (\xSharp)_v
\]
is the original expression minimised by both the LP and ILP, where as said
in \ref{continuousrules} the
LP solution is at least as good as the ILP one (which is however the actual optimal cost
for the vertex cover problem). So we have
\[
  \cost(\covI)
   \leq \cost(\cov)
   \leq 2\cdot \sum_{v\in V}c_v \cdot (\xSharp)_v
   \leq 2\cdot \sum_{v\in V}c_v \cdot (\xSharpI)_v
   = 2\cdot \cost(\covI).
\]

\subex{}
\newcommand{\centcost}{\mathscr{V}}
We consider star-shaped graphs in which the central vertex has a cost of $\centcost$
and each of the $N$ satellite vertices has cost $1$. Clearly, the optimum vertex
cover for this scenario is to cover only the central vertex if $\centcost<N$ (giving
a total cost $\centcost$), and to cover all satellites but not the center otherwise, which
has a cost of $N$. We use a numerical LP-solver to check how well the relaxated version
of the problem matches this.

\simplegraphics{ex4/starcosts.pdf}
   {Numerically solving the LP-problem shows that, for star-shaped graphs, the relaxated
    solution perfectly matches the exact integral one.}





